---
title: "Specification of the 'tiny_labelled' class"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Specification of the 'tiny_labelled' class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinylabels)
```

# Setting a variable label

```{r}
x <- 1:4
variable_label(x) <- "A variable label"
```

# Extracting and replacing parts of an object

If elements of a vector are extracted, the new vector retains label and class `tiny_labelled`.

```{r}
x[1:2]
str(x)
x[1:2] <- 5:6
str(x)
```

# Math, Ops, Summary, Complex

If a vector is modified (e.g., via mathematical operations), label and `tiny_labelled` class are removed.

```{r}
str(exp(x))
```

```{r}
str(x + 1)
```

```{r}
str(min(x))
```

```{r}
str(Re(x))
```

# Coercion

Vectors of class `tiny_labelled` keep label and class if they are modified via `as.character()`, `as.numeric()`, etc.

```{r}
as.character(x)
```

If `keep_label = FALSE`, label and class are removed.

```{r}
as.character(x, keep_label = FALSE)
```

`as()` methods are always strict, i.e. label and class are removed:

```{r}
as(x, "character")
```


# Data-frame methods

It is possible to assign multiple labels to the columns of a data frame by passing a named list or a named vector of
key-value pairs. Note that mixing different types of labels (e.g. `character` and `expression` labels) is only possible
if the right-hand side is a list.

```{r}
variable_label(npk) <- c(
  N = "Nitrogen"
  , P = "Phosphate"
)
variable_label(npk) <- list(
  yield = expression(bar(Yield))
)
variable_label(npk)
```

