---
title: "Introduction to Tiny Variable Labels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Tiny Variable Labels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinylabels)
```

# Philosophy

- lightweight
    - no imports, good choice for package developers
- minimum side effects with existing functions
    - haven
    - Hmisc
- Useful error and warning messages

# Basic Usage

```{r}
x <- rnorm(6)
variable_label(x) <- "Values randomly drawn from a standard-normal distribution"
x
```
```{r}
# Extract the variable label from a vector (e.g., a numeric vector)
variable_label(x)
```


It is also possible to assign variable labels to (all or a subset of) the columns of a data frame.
For instance, consider the built-in data set `npk`:

```{r}
# View original data set ----
head(npk)
```



```{r}
# Assign labels to the built-in data set 'npk' ----
variable_label(npk) <- c(
  N = "Nitrogen"
  , P = "Phosphate"
  , yield = "Pea yield"
)

# Extract the variable labels ----
variable_label(npk)
```
The return value of `variable_label()` is a named list, where each list element corresponds to a column of a data frame.

To assign labels to the columns of a data frame, the right-hand side of the assignment has to be a **named** vector or list.
When trying to set columns that are not present in the data frame, a meaningful error  message is thrown:

```{r error = TRUE}
variable_label(npk) <- c(wrong_column_name = "A supposedly terrific label")
```

If you want to remove labels (and the corresponding S3 class) from a vector or all columns of a data frame, you may use function `unlabel`:

```{r}
# Return as a simple factor ----
unlabel(npk$N)

# Remove all labels (and class 'tiny_labelled') from all columns ----
npk <- unlabel(npk)
head(npk)
```



# Supporting the *tidyverse*

```{r}
data_1 <- data_2 <- data.frame(
  x = rnorm(10)
  , y = rnorm(10)
)

variable_label(data_1) <- c(x = "Label for x", y = "Label for y")
variable_label(data_2) <- c(x = "Label for x", y = "Another label for y")

library(dplyr)
combined_data <- bind_rows(data_1, data_2)
variable_label(combined_data)
```

